Liveness-Based Garbage Collection for Lazy Languages

### 2015-07-11 paper/paper.tex
table split.

\subsection{Liveness-based collector with reachability-based closure
  collection}
\begin{algorithm}[t]
  \Fn{\Lgc{}}
     {
       \For {each reference $\mathit{ref}$ in root set}
            {\LCopy($\mathit{ref}$)\;}
            $\mathit{copyReferencesOnPrintStack()}$\;  
     }
     \Fn{\LCopy{$\mathit{ref}$}}
        {
          \eIf{cellType$(\mathit{ref})$ is $cons$}
              {
                \eIf {$\mathit{ref}$ is found live using dfa and has not been copied using
                  reachability}
                     {
                       $\mathit{newRef}$ =  $\mathit{dupHeapcell(ref})$\; 
                       $\mathit{ newCar}  =$  \LCopy$\mathit{(ref.car)}$\;
                       $\mathit{ newRef.car}   = \mathit{ newCar}$\;
                       $\mathit{ newCdr}  =$  \LCopy$\mathit{(ref.cdr)}$\;
                       $\mathit{ newRef.cdr}   = \mathit{newCdr}$\;
                     }
                     { $\mathit{makenull(ref)}$\;}}
              {
                \If {cellType$\mathit{(ref)}$ is closure and 
                  has not been copied}
                    { 
                      $\mathit{newRef}$  = \RCopy$(\mathit{ref})$\;
                    }
              }
        }
        \caption{LGC with reachability-based closure collection.\label{algo:lgc-b}}
\end{algorithm}
\begin{table*}[ht!]
\caption{Experimental results comparing RGC and LGC}
\label{tab:experimental-results}
\centering
\include{autotable}
\end{table*}

@@

\begin{table*}[ht!]
\caption{Comparing LGC times for different strategies}
\label{tab:compare-live-clo-reach-clo}
\centering
\begin{tabular}
{| l | r | r | r | r |}
\hline
Prog Name & Heap size & Num GC & Time taken using reachability for closure  & Time taken using liveness for closure \\ \hline
\verb@nqueens(8)@ & 10000 & 183 & 0.0087s & 0.2267s\\ \hline
\verb@lambda@&10000&11&0.0010s&4.9922s\\ \hline
\verb@lcss@&10000&1&0.00008s&34.6684s\\ \hline
\end{tabular}
\end{table*}
To compare the  effect of using reachability to  copy closures instead
of liveness we take some representative programs from our benchmark suite and
execute       them      with       both.        As      shown       in
Table~\ref{tab:compare-live-clo-reach-clo},  the   number  of  garbage
collections in both cases are same but copying closures using liveness
takes a lot more time compared to reachability based copying.

>>>>

\subsection{Liveness-based collector with reachability-based closure
  collection}
\begin{algorithm}[t]
  \Fn{\Lgc{}}
     {
       \For {each reference $\mathit{ref}$ in root set}
            {\LCopy($\mathit{ref}$)\;}
            $\mathit{copyReferencesOnPrintStack()}$\;  
     }
     \Fn{\LCopy{$\mathit{ref}$}}
        {
          \eIf{cellType$(\mathit{ref})$ is $cons$}
              {
                \eIf {$\mathit{ref}$ is found live using dfa and has not been copied using
                  reachability}
                     {
                       $\mathit{newRef}$ =  $\mathit{dupHeapcell(ref})$\; 
                       $\mathit{ newCar}  =$  \LCopy$\mathit{(ref.car)}$\;
                       $\mathit{ newRef.car}   = \mathit{ newCar}$\;
                       $\mathit{ newCdr}  =$  \LCopy$\mathit{(ref.cdr)}$\;
                       $\mathit{ newRef.cdr}   = \mathit{newCdr}$\;
                     }
                     { $\mathit{makenull(ref)}$\;}}
              {
                \If {cellType$\mathit{(ref)}$ is closure and 
                  has not been copied}
                    { 
                      $\mathit{newRef}$  = \RCopy$(\mathit{ref})$\;
                    }
              }
        }
        \caption{LGC with reachability-based closure collection.\label{algo:lgc-b}}
\end{algorithm}

\begin{table*}[ht!]
\caption{Comparing LGC times for different strategies}
\label{tab:compare-live-clo-reach-clo}
\centering
\begin{tabular}
{| l | r | r | r | r |}
\hline
Prog Name & Heap size & Num GC & Time taken using reachability for closure  & Time taken using liveness for closure \\ \hline
\verb@nqueens (reduced) @ & 10000 & 183 & 0.0087s & 0.2267s\\ \hline
\verb@lambda@&10000&11&0.0010s&4.9922s\\ \hline
\verb@lcss@&10000&1&0.00008s&34.6684s\\ \hline
\end{tabular}
\end{table*}

\begin{table*}[ht!]
\caption{Compile time and Garbage Collection statistics}
\label{tab:exp-results}
\centering
 \begin{tabular}{|l||c@{\ }|@{\ }c@{\ }|@{\ }c@{\ }|@{\ }c@{\ }|@{\ }c@{\ }|@{\ }c@{\ }|@{\ }c@{\ }|@{\ }c@{\ }|@{\ }c|} \hline
   \multicolumn{1}{|r||}{Program} & \verb@lambda@ & \verb@nperm@ & \verb@treejoin@ & \verb@lcss@ & \verb@sudoku@ & \verb@fibheap@ & \verb@nqueens@ & \verb@knightstour@ & \verb@gc_bench@ \\ \hline\hline
   \#CFG Nonterminals            & 1078          & 794          & 709             & 839         & 1758          & 780            & 529            & 659                & 417             \\ \hline
   \#CFG Rules                   & 1823          & 1038         & 1610            & 1752        & 2509          & 1358           & 748            & 909                & 486             \\  \hline 
   \#DFA States                  & 3312          & 2055         & 2332            & 2340        & 5672          & 2254           & 1229           & 1742               & 729             \\\hline
   \#DFA Transitions             & 5792          & 3632         & 4235            & 4261        & 10857         & 4076           & 2078           & 3101               & 1124            \\\hline
   DFA Generation Time (sec)      & 14.7          & 1.1          & 2690.5          & 13.3        & 798.5         & 49.8           & 0.5            & 5.1                & 0.1             \\\hline
 \end{tabular}\\
 (a) Data for Compile-time Liveness Analysis \\[3pt]
\include{autotable}
\end{table*}

@@

To compare the  effect of using reachability to  copy closures instead
of liveness we take some representative programs from our benchmark suite and
execute       them      with       both.        As      shown       in
Table~\ref{tab:compare-live-clo-reach-clo},  the   number  of  garbage
collections in both cases are same but copying closures using liveness
takes a lot more time compared to reachability based copying.
